Cycles of the logistic map

### 2013-04-15 doc/pc-ijbc.tex
doc: change \phi to \varphi

\begin{remark}
%\begin{remark}[1]
  For complex $R$ and $\vx$,
  $\lambda$ should be generalized from $\pm1$
  to any $\lambda = \exp(i\phi)$,
  where $\phi \in [0, 2\pi]$,
  the algorithm still applies.
  By increasing $\phi$ from 0 to $2\pi$,
  we can trace a two-dimensional region of a complex $R$
  for stable cycles.
  These regions are bulbs in the Mandelbrot set,
  %(however, we need the replacement
  %$x_k \rightarrow -z_k$
  %and
  %$R \rightarrow -c$)
  see ref. \cite{stephenson} and \reffig{logbulb}.

@@

\begin{table}[h]\footnotesize
  %\caption{
  \tbl{
  Characteristic polynomials $A_n\RX$ of
  the fixed points of $f^n$ of
  the simplified logistic map \refeq{logmaps}.
  }
%\begin{center}
{
\begin{tabular}{lc}
\hline
  $n$
& $A_n(R, \; \lambda = 2^n X)
  \big\slash
   2^{n \NB(n)}$ $^\dagger$
\\
\hline
1
&
$X^2-X-R$
\\
2
&
  $
  \big[ \cancel{(R-X)^2 - X}\big]
  (R + X - 1)
  $
\\
3
&
$\begin{aligned}
\big[\cancel{
  X^2-(3R+1)X-R^3
}\big]
  \big[
  (X-1)^2
 + X R
 - R(R-1)^2
  \big]
\end{aligned}$
\\
4
&
$\begin{aligned}
\big[ \cancel{
  (R-1)^2 - X
} \big]
\big[ \cancel{
(R^2+X)^2 -(2R+1)^2 X
} \big]
\big[
  R^6-3 R^5+(X+3) (R^4 - R^3) -(X+2)(X-1) R^2 - (X-1)^3
\big]
\end{aligned}$
\\
5
&
\begin{minipage}{.97\linewidth}
\vspace*{2mm}
$\begin{aligned}
&\big[ \cancel{
  X^2-5 X (R^2+R)-X - R^5
  %R^5+5 X (R^2+R)-X^2+X
} \big]
\big[
-R^{15}+8 R^{14}-28 R^{13}+(X+60) R^{12}-(7 X+94) R^{11}
+(3 X^2 +20 X
\\
+&116) R^{10}
-(11 X^2+33 X+114) R^9
+2(3 X^2+20 X+47) R^8
-(2 X^3-20 X^2+37 X+69) R^7 \\
+&(3 X-11)(3X^2-3X+44) R^6
%-(3 X^4+17 X^3-53 X^2+7 X+26) R^5
-(X-1) (3 X^3+20 X^2-33 X-26) R^5
+(X-1)^2 (3 X^2+27 X+14) R^4 \\
+&(X-1)^3 (6 X+5) R^3
+(X-1)^4 (X+2) R^2+(X-1)^5 R+(X-1)^6
\big]
\end{aligned}$
\vspace{1mm}
\end{minipage}
\\
6
&
$\begin{aligned}
&\big[\cancel {
  (R-1)^3 + X
}\big]
\big[\cancel {
 (R^3 - X)^2 - (3R+1)^2 X
}\big]
\big[\cancel {
\big(R(R-1)^2-1-X\big)^2-(R-2)^2 X
}\big]
\big[
R^{27}-13 R^{26}+78 R^{25}\\
&+(X-293) R^{24}
+\dots
+(X-1)^6 (X^2+10 X+3) R^3
+(X-1)^7 (X+1) R^2-(X-1)^8 R+(X-1)^9
\big]
\end{aligned}$
\\
7
&
\begin{minipage}{.97\linewidth}
\vspace{1mm}
$\begin{aligned}
&\big[\cancel {
  -R^7 - 7 X R (R+1)^2 + X^2 - X
}\big]
\big[
-R^{63}+32 R^{62}-496 R^{61}
+4976 R^{60} + (X - 36440) R^{59}
- (30X
\\
-&208336) R^{58}
+\dots
+(X-1)^{15} (X^2+14 X+5) R^3
+2(X-1)^{16} (X+1) R^2
+(X-1)^{17} R
+(X-1)^{18}
\big]
\end{aligned}$
\end{minipage}
\\
\vdots & \vdots
\\
\hline
\multicolumn{2}{p{\textwidth}}{
$^\dagger$
$\NB(n) = (1/n) \sum_{d|n} \phi(n/d) 2^d$ [\refeq{necklace}]
is the number of the square-free cyclic polynomials.
%
The change of variable $\lambda \rightarrow X$ and
  the division by $2^{n \NB(n)}$ make the polynomials more compact.
%
The irrelevant factors from shorter cycles
(see \refsec{primfac}) are struck out.
The polynomials of $r$ for the original logistic map \refeq{logmap}
can be obtained by $R\rightarrow r(r-2)/4$.
} \\
\hline
\end{tabular}
\label{tab:Anlog}
}
%\end{center}
\end{table}

The algorithm was coded into a Mathematica program,
which was used to compute the polynomials for $n$ up to 13.
The polynomials for a general $\lambda$
and those at $\lambda = \pm1$ (onset and bifurcation points) are listed
in \reftab{Anlog} and \reftab{Pnlog}, respectively, for some small $n$.
%
For complex $R$, $\lambda$, and $\vx$,
the method can also compute the region of stability for $R$,
with $\lambda$ being $\exp(i\phi)$ \big[$\phi \in (0, 2\pi)$\big]
  instead of $\pm1$;
the results are shown in \reffig{logbulb}
for $n$ up to $8$.
%
For polynomials of larger $n$, see the website in \refsec{end}.
The representative $r$ values are listed in \reftab{rval}.

@@

\begin{figure}[h]
  \begin{minipage}{\linewidth}
    \begin{center}
        \includegraphics[angle=-90, width=.67\linewidth]{logbulb.pdf}
    \end{center}
  \end{minipage}%
  \caption{
  \label{fig:logbulb}
  Stable regions of the $n$-cycles of the simplified logistic map \refeq{logmaps}
  with a complex $R$;
  obtained from $P_n(R, e^{i\phi}) = 0$ with $\phi \in [0, 2\pi]$,
  cf. Fig. 3 in \cite{stephenson}.
}
\end{figure}

@@

\begin{table}[h]\footnotesize
%\caption{
\tbl{
Two ways of counting the $N(4) = 6$ necklaces for $n = 4$.
}
%\begin{center}
{
%\begin{tabularx}{\textwidth}{c | c  c | c | c | c | c | c }
\begin{tabular}{c | c  c | c | c | c | c | c }
\hline
        & \multicolumn{2}{c|}{$m = 1$}
        & $m = 2$
        & $m = 4$   & $2^d$ & $\phi(n/d)$ & $\phi(n/d) 2^d$\\
\hline
$d = 1$ & \hspace{1mm} $\mathbf{0}_{\times4}$ \hspace{1mm}
        & \hspace{1mm} $\mathbf{1}_{\times4}$ \hspace{1mm} & & & $2 = 2$ & $2$ & 4 \\
\hline
$d = 2$ & $\mathbf{00}_{\times2}$ & $\mathbf{11}_{\times2}$
        & \hspace{1mm} $\mathbf{10}_{\times2} (01_{\times2})$
          \hspace{1mm} & & $2 + 2 =4$ & $1$ & 4 \\
\hline
\multirow{3}{*}{$d = 4$}
      & & & & \hspace{1mm} $\mathbf{1000} (0100, 0010, 0001)$
              \hspace{1mm} & $2 + 2 + 12$& \\
      & $\mathbf{0000}$ & $\mathbf{1111}$ & $\mathbf{1010} (0101)$
            & $\mathbf{1100} (1001, 0110, 0011)$ & $=16$ & $1$ & 16 \\
      & & & &   $\mathbf{1110} (1101, 1011, 0111)$ & &  \\
\hline
$m \sum_{m|d, d|n} \phi(\frac{n}{d})$
      & \multicolumn{2}{c|}{$1\cdot(2+1+1) = 4$}
      & $2\cdot(1+1) = 4$ & $4\cdot1 = 4$ &  \multicolumn{2}{c|}{} &  $\downarrow$\\
\hline
$N(n)\cdot n$ & \multicolumn{2}{c|}{$2 \cdot 4$} & $1 \cdot 4$ & $3 \cdot 4$ &
  \multicolumn{2}{c|}{$\rightarrow$} & $6\cdot4 = 24$ \\
\hline
\multicolumn{8}{p{\linewidth}}{
  Bold strings are necklaces; others are their cyclic versions.
}
\\
\multicolumn{8}{p{\linewidth}}{
The subscript of a binary string means the number of repeats;
e.g.,
$\mathbf{0}_{\times4}$ means $\mathbf{0}$ repeated four times, or $\mathbf{0000}$;
$\mathbf{10}_{\times2}$ means $\mathbf{10}$ repeated twice, or $\mathbf{1010}$, etc.
}
\\
\hline
\end{tabular}
%\end{tabularx}
%\end{center}
\label{tab:countnecklace}
}
\end{table}

To compute the number of the binary necklaces $N(n)$,
we construct a sum for the length-$n$ binary strings and count it in two ways.
\reftab{countnecklace} shows the example for the $n=4$ case.
%
For each divisor $d$ of $n$, we collect
  all length-$n$ binary strings
  whose periods $m$ divide $d$.
%
The total is $2^d$,
  for we have enumerated all binary strings whose periods divide $d$.
%and each period-$m$ necklace contributes $m$ cyclic versions.
%
We then weight them by the Euler's totient function $\phi(n/d)$.
%
Here, $\phi(m)$ gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\phi(1) = 1$,
  $\phi(3) = 2$ for 1 and 2,
  $\phi(6) = 2$ for 1 and 5.
We repeat the process over other divisors $d$ of $n$,
  and the resulting sum is $\sum_{d|n} \phi(n/d) 2^d$.
%
The process for a fixed $d$ is exemplified
  by a row in \reftab{countnecklace}.

@@

We can count the above sum in another way.
We recall that a period-$m$ necklace always contributes
  $m$ strings in the above process for a fixed $d$,
  and it does so for all multiples $d$ of $m$.
So the total weighted contribution by this necklace is
\begin{equation}
m \sum_{m|d, d|n} \phi(n/d) =
 m \sum_{\frac{n}{d}|\frac{n}{m}} \phi(n/d) = m \,  \frac{n}{m}  = n,
\label{eq:mphind}
\end{equation}
where we have used the identity $\sum_{d'|m'} \phi(d') = m'$,
with $d' = n/d$ and $m' = n/m$.
%
Summing over the necklaces yields $N(n) \, n$. Thus,
$N(n) = (1/n) \sum_{d|n} \phi(n/d) 2^d$.
The process for a fixed necklace is exemplified
  by a column in \reftab{countnecklace}.
%

@@

\begin{theorem}
The number $\NB(n)$ of the square-free cyclic polynomials $C_p(\vx)$ ($p \in \Bset$)
  formed by $\vx = \{x_1, \ldots, x_n\}$ is
\begin{equation}
  \NB(n) = N(n) = (1/n) \sum_{d|n} \phi(n/d) 2^{\,d},
\label{eq:necklace}
\end{equation}
%
which is also equal to $\deg_\lambda A_n\RX$.
\label{thm:necklace}
\end{theorem}
%
$N(n)$ = 2, 3, 4, 6, 8, 14, 20, 36, 60, 108, 188, 352, 632, \ldots,
  starting from $n = 1$.

@@

Since the period $d$ of a length-$n$ necklace always divides $n$,
we have
\begin{equation}
N(n) = \sum_{d|n} L(d).
\label{eq:necklacelyndon}
\end{equation}
%
We can also show this by explicit computation:
\begin{align*}
\sum_{d | n} L(d)
  &= \sum_{d | n} (1/d) \sum_{c |  d} \mu(d/c) 2^{\,c}
   = (1/n) \sum_{c | n} 2^{\,c} \sum_{d'  |  (n/c)}
        \mu\left(\frac{n/c}{d'}\right) d' \\
  &= (1/n) \sum_{c \, | \, n} 2^{\,c} \, \phi(n/c)
  = N(n),
\end{align*}
where we have used $\phi(m) = \sum_{d' | m} \mu(\frac{m}{d'}) \, d'$,
which is the inversion of $m = \sum_{d' | m} \phi(d')$,
and \refeq{necklace}.
%
We will use \refeq{necklacelyndon} in proving the next theorem.

@@

\begin{theorem}
The degrees in $R$ of $A_n\RX$, $P_n\RX$ and $S_n(R)$ are
\begin{subequations}
\begin{align}
\deg_R A_n\RX &= \sum_{d|n} \phi(n/d) 2^{d-1}, \\
\deg_R P_n\RX &= \sum_{d|n} \mu(n/d) 2^{d-1} \equiv \beta(n), \\
\deg S_n(R) &= \beta(n) - \sum_{d | n, d < n} \beta(d) \, \phi(n/d).
\end{align}
\label{eq:degR}
\end{subequations}
\label{thm:degR}
\end{theorem}

@@

\begin{figure}[h]
  \begin{minipage}{\linewidth}
    \begin{center}
        \includegraphics[angle=-90, width=\linewidth]{cubbulb.pdf}
    \end{center}
  \end{minipage}%
  \caption{
  \label{fig:cubbulb}
  Stable regions of the $n$-cycles of the cubic map \refeq{cubic}
  with a complex $R$;
  obtained from $P_n(R, e^{i\phi}) = 0$ with $\phi \in [0, 2\pi]$
  cf. \reffig{logbulb}.
}
\end{figure}

@@

%Now, the number of the cube-free cyclic polynomials
%is now equal to that of total necklaces
%\[
%  N_3(n) = (1/n) \sum_{d|n} \phi(n/d) 3^d.
%\]
%
%The proof is similar to that in \refsec{degprimfac}.
%For even necklaces, we have

\begin{theorem}
The number of the even ternary necklaces
or the cube-free cyclic polynomials for the cubic map of even degrees in $\vx$
is given by
\begin{equation}
  N_e(n) = \frac{1}{n} \sum_{c d = n} \phi(c)
    \left[
      3^d - \odd(c) \frac{3^d-1}{2}
    \right],
\label{eq:cubnecklace}
\end{equation}
where
$\odd(c) \equiv \big[1 - (-1)^c\big]/2$ is 1 if $c$ is odd or 0 if even.
\label{thm:cubnecklace}
\end{theorem}

@@

\begin{table}[h]\footnotesize
\tbl{
%\caption{
Two ways of counting the $N_e(2) = 4$ ternary necklaces for $n = 2$.
}
%\begin{center}
{
\begin{tabular}{c | c  c c | c  | c | c | c | c }
%\begin{tabularx}{\textwidth}{c | c  c c | c  | c | c | c | c }
\hline
        & \multicolumn{3}{c|}{$m = 1$}
        & $m = 2$
        & $c = n/d$
        & $T(d, c)$ $^\dagger$
        & $\phi(c)$ & $\phi(n/d) T(d, c)$\\
\hline
$d = 1$ & \hspace{1mm} $\mathbf{0}_{\times2}$ \hspace{1mm}
        & \hspace{1mm} $\mathbf{1}_{\times2}$ \hspace{1mm}
        & \hspace{1mm} $\mathbf{2}_{\times2}$ \hspace{1mm}
        & & $2$ & $3$ & $1$ & 3 \\
\hline
$d = 2$ & $\mathbf{00}$
        & $\mathbf{11}$
        & $\mathbf{22}$
        & \hspace{1mm} $\mathbf{20} (02)$ $^\ddagger$ \hspace{1mm}
        & $1$ & $3+1\cdot2=5$ & $1$ & 5 \\
\hline
$m \sum_{m|d, d|n} \phi(\frac{n}{d})$
      & \multicolumn{3}{c|}{$1\cdot(1+1) = 2$}
      & $2\cdot1 = 2$ &    \multicolumn{3}{c|}{} &  $\downarrow$\\
\hline
$N_e(n)\cdot n$
    & \multicolumn{3}{c|}{$3 \cdot 2$} & $1 \cdot 2$ &
  \multicolumn{3}{c|}{$\rightarrow$} & $4\cdot2 = 3 + 5$ \\
\hline
\multicolumn{9}{p{\linewidth}}{
  Bold strings are necklaces; others are their cyclic versions.
}
\\
\multicolumn{9}{p{\linewidth}}{
The subscript of a string means the number of repeats;
e.g.,
$\mathbf{1}_{\times2}$ means $\mathbf{1}$ repeated twice, or $\mathbf{11}$;
}
\\
\multicolumn{9}{p{\linewidth}}{
$^\dagger$ $T(d, c) \equiv 3^d - \odd(c)(3^d-1)/2$,
which is $3^d$ if $c$ is even,
or $(3^d+1)/2$ if $c$ is odd.
}
\\
\multicolumn{9}{p{\linewidth}}{
$^\ddagger$
The odd binary strings $\mathbf{10}$, $01$, $\mathbf{21}$, and $12$
  do not contribute to the sum, and are excluded from $T(d, c)$.
}
\\
\hline
\end{tabular}
%\end{tabularx}
%\end{center}
\label{tab:countcubnecklace}
}
\end{table}

The rest counting process is similar to that in \refsec{degprimfac}:
%
we construct the sum $n N_e(n)$ by two ways, as exemplified in \reftab{countcubnecklace}.
%
In the first way,
  for a fixed $d$ ($d|n$),
  if $c\equiv n/d$ is even,
    we count all ternary strings whose period $m$ divide $d$,
  but if $c$ is odd, we count only ternary strings
    whose first $d$ beads are even
    (because repeating an odd string an odd number times
    does not yield an even string);
  in either case, we multiple the result by $\phi(c)$.
%
The process for a fixed $d$ corresponds to a row of \reftab{countcubnecklace}.
%
Repeating the process for all $d$ gives
  $\sum_{cd = n} T(d, c) \, \phi(c)$,
  where $T(d, c)$ is the total number $3^d$ of period-$d$ ternary strings if $c$ is even,
  or the number $(3^d+1)/2$ of even strings if $c$ is odd.

>>>>

\begin{remark}
%\begin{remark}[1]
  For complex $R$ and $\vx$,
  $\lambda$ should be generalized from $\pm1$
  to any $\lambda = \exp(i\theta)$,
  where $\theta \in [0, 2\pi]$,
  the algorithm still applies.
  By increasing $\theta$ from 0 to $2\pi$,
  we can trace a two-dimensional region of a complex $R$
  for stable cycles.
  These regions are bulbs in the Mandelbrot set,
  %(however, we need the replacement
  %$x_k \rightarrow -z_k$
  %and
  %$R \rightarrow -c$)
  see ref. \cite{stephenson} and \reffig{logbulb}.

@@

\begin{table}[h]\footnotesize
  %\caption{
  \tbl{
  Characteristic polynomials $A_n\RX$ of
  the fixed points of $f^n$ of
  the simplified logistic map \refeq{logmaps}.
  }
%\begin{center}
{
\begin{tabular}{lc}
\hline
  $n$
& $A_n(R, \; \lambda = 2^n X)
  \big\slash
   2^{n \NB(n)}$ $^\dagger$
\\
\hline
1
&
$X^2-X-R$
\\
2
&
  $
  \big[ \cancel{(R-X)^2 - X}\big]
  (R + X - 1)
  $
\\
3
&
$\begin{aligned}
\big[\cancel{
  X^2-(3R+1)X-R^3
}\big]
  \big[
  (X-1)^2
 + X R
 - R(R-1)^2
  \big]
\end{aligned}$
\\
4
&
$\begin{aligned}
\big[ \cancel{
  (R-1)^2 - X
} \big]
\big[ \cancel{
(R^2+X)^2 -(2R+1)^2 X
} \big]
\big[
  R^6-3 R^5+(X+3) (R^4 - R^3) -(X+2)(X-1) R^2 - (X-1)^3
\big]
\end{aligned}$
\\
5
&
\begin{minipage}{.97\linewidth}
\vspace*{2mm}
$\begin{aligned}
&\big[ \cancel{
  X^2-5 X (R^2+R)-X - R^5
  %R^5+5 X (R^2+R)-X^2+X
} \big]
\big[
-R^{15}+8 R^{14}-28 R^{13}+(X+60) R^{12}-(7 X+94) R^{11}
+(3 X^2 +20 X
\\
+&116) R^{10}
-(11 X^2+33 X+114) R^9
+2(3 X^2+20 X+47) R^8
-(2 X^3-20 X^2+37 X+69) R^7 \\
+&(3 X-11)(3X^2-3X+44) R^6
%-(3 X^4+17 X^3-53 X^2+7 X+26) R^5
-(X-1) (3 X^3+20 X^2-33 X-26) R^5
+(X-1)^2 (3 X^2+27 X+14) R^4 \\
+&(X-1)^3 (6 X+5) R^3
+(X-1)^4 (X+2) R^2+(X-1)^5 R+(X-1)^6
\big]
\end{aligned}$
\vspace{1mm}
\end{minipage}
\\
6
&
$\begin{aligned}
&\big[\cancel {
  (R-1)^3 + X
}\big]
\big[\cancel {
 (R^3 - X)^2 - (3R+1)^2 X
}\big]
\big[\cancel {
\big(R(R-1)^2-1-X\big)^2-(R-2)^2 X
}\big]
\big[
R^{27}-13 R^{26}+78 R^{25}\\
&+(X-293) R^{24}
+\dots
+(X-1)^6 (X^2+10 X+3) R^3
+(X-1)^7 (X+1) R^2-(X-1)^8 R+(X-1)^9
\big]
\end{aligned}$
\\
7
&
\begin{minipage}{.97\linewidth}
\vspace{1mm}
$\begin{aligned}
&\big[\cancel {
  -R^7 - 7 X R (R+1)^2 + X^2 - X
}\big]
\big[
-R^{63}+32 R^{62}-496 R^{61}
+4976 R^{60} + (X - 36440) R^{59}
- (30X
\\
-&208336) R^{58}
+\dots
+(X-1)^{15} (X^2+14 X+5) R^3
+2(X-1)^{16} (X+1) R^2
+(X-1)^{17} R
+(X-1)^{18}
\big]
\end{aligned}$
\end{minipage}
\\
\vdots & \vdots
\\
\hline
\multicolumn{2}{p{\textwidth}}{
$^\dagger$
$\NB(n) = (1/n) \sum_{d|n} \varphi(n/d) 2^d$ [\refeq{necklace}]
is the number of the square-free cyclic polynomials.
%
The change of variable $\lambda \rightarrow X$ and
  the division by $2^{n \NB(n)}$ make the polynomials more compact.
%
The irrelevant factors from shorter cycles
(see \refsec{primfac}) are struck out.
The polynomials of $r$ for the original logistic map \refeq{logmap}
can be obtained by $R\rightarrow r(r-2)/4$.
} \\
\hline
\end{tabular}
\label{tab:Anlog}
}
%\end{center}
\end{table}

The algorithm was coded into a Mathematica program,
which was used to compute the polynomials for $n$ up to 13.
The polynomials for a general $\lambda$
and those at $\lambda = \pm1$ (onset and bifurcation points) are listed
in \reftab{Anlog} and \reftab{Pnlog}, respectively, for some small $n$.
%
For complex $R$, $\lambda$, and $\vx$,
the method can also compute the region of stability for $R$,
with $\lambda$ being $\exp(i\theta)$ \big[$\theta \in (0, 2\pi)$\big]
  instead of $\pm1$;
the results are shown in \reffig{logbulb}
for $n$ up to $8$.
%
For polynomials of larger $n$, see the website in \refsec{end}.
The representative $r$ values are listed in \reftab{rval}.

@@

\begin{figure}[h]
  \begin{minipage}{\linewidth}
    \begin{center}
        \includegraphics[angle=-90, width=.67\linewidth]{logbulb.pdf}
    \end{center}
  \end{minipage}%
  \caption{
  \label{fig:logbulb}
  Stable regions of the $n$-cycles of the simplified logistic map \refeq{logmaps}
  with a complex $R$;
  obtained from $P_n(R, e^{i\theta}) = 0$ with $\theta \in [0, 2\pi]$,
  cf. Fig. 3 in \cite{stephenson}.
}
\end{figure}

@@

\begin{table}[h]\footnotesize
%\caption{
\tbl{
Two ways of counting the $N(4) = 6$ necklaces for $n = 4$.
}
%\begin{center}
{
%\begin{tabularx}{\textwidth}{c | c  c | c | c | c | c | c }
\begin{tabular}{c | c  c | c | c | c | c | c }
\hline
        & \multicolumn{2}{c|}{$m = 1$}
        & $m = 2$
        & $m = 4$   & $2^d$ & $\varphi(n/d)$ & $\varphi(n/d) 2^d$\\
\hline
$d = 1$ & \hspace{1mm} $\mathbf{0}_{\times4}$ \hspace{1mm}
        & \hspace{1mm} $\mathbf{1}_{\times4}$ \hspace{1mm} & & & $2 = 2$ & $2$ & 4 \\
\hline
$d = 2$ & $\mathbf{00}_{\times2}$ & $\mathbf{11}_{\times2}$
        & \hspace{1mm} $\mathbf{10}_{\times2} (01_{\times2})$
          \hspace{1mm} & & $2 + 2 =4$ & $1$ & 4 \\
\hline
\multirow{3}{*}{$d = 4$}
      & & & & \hspace{1mm} $\mathbf{1000} (0100, 0010, 0001)$
              \hspace{1mm} & $2 + 2 + 12$& \\
      & $\mathbf{0000}$ & $\mathbf{1111}$ & $\mathbf{1010} (0101)$
            & $\mathbf{1100} (1001, 0110, 0011)$ & $=16$ & $1$ & 16 \\
      & & & &   $\mathbf{1110} (1101, 1011, 0111)$ & &  \\
\hline
$m \sum_{m|d, d|n} \varphi(\frac{n}{d})$
      & \multicolumn{2}{c|}{$1\cdot(2+1+1) = 4$}
      & $2\cdot(1+1) = 4$ & $4\cdot1 = 4$ &  \multicolumn{2}{c|}{} &  $\downarrow$\\
\hline
$N(n)\cdot n$ & \multicolumn{2}{c|}{$2 \cdot 4$} & $1 \cdot 4$ & $3 \cdot 4$ &
  \multicolumn{2}{c|}{$\rightarrow$} & $6\cdot4 = 24$ \\
\hline
\multicolumn{8}{p{\linewidth}}{
  Bold strings are necklaces; others are their cyclic versions.
}
\\
\multicolumn{8}{p{\linewidth}}{
The subscript of a binary string means the number of repeats;
e.g.,
$\mathbf{0}_{\times4}$ means $\mathbf{0}$ repeated four times, or $\mathbf{0000}$;
$\mathbf{10}_{\times2}$ means $\mathbf{10}$ repeated twice, or $\mathbf{1010}$, etc.
}
\\
\hline
\end{tabular}
%\end{tabularx}
%\end{center}
\label{tab:countnecklace}
}
\end{table}

To compute the number of the binary necklaces $N(n)$,
we construct a sum for the length-$n$ binary strings and count it in two ways.
\reftab{countnecklace} shows the example for the $n=4$ case.
%
For each divisor $d$ of $n$, we collect
  all length-$n$ binary strings
  whose periods $m$ divide $d$.
%
The total is $2^d$,
  for we have enumerated all binary strings whose periods divide $d$.
%and each period-$m$ necklace contributes $m$ cyclic versions.
%
We then weight them by the Euler's totient function $\varphi(n/d)$.
%
Here, $\varphi(m)$ gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2,
  $\varphi(6) = 2$ for 1 and 5.
We repeat the process over other divisors $d$ of $n$,
  and the resulting sum is $\sum_{d|n} \varphi(n/d) 2^d$.
%
The process for a fixed $d$ is exemplified
  by a row in \reftab{countnecklace}.

@@

We can count the above sum in another way.
We recall that a period-$m$ necklace always contributes
  $m$ strings in the above process for a fixed $d$,
  and it does so for all multiples $d$ of $m$.
So the total weighted contribution by this necklace is
\begin{equation}
m \sum_{m|d, d|n} \varphi(n/d) =
 m \sum_{\frac{n}{d}|\frac{n}{m}} \varphi(n/d) = m \,  \frac{n}{m}  = n,
\label{eq:mphind}
\end{equation}
where we have used the identity $\sum_{d'|m'} \varphi(d') = m'$,
with $d' = n/d$ and $m' = n/m$.
%
Summing over the necklaces yields $N(n) \, n$. Thus,
$N(n) = (1/n) \sum_{d|n} \varphi(n/d) 2^d$.
The process for a fixed necklace is exemplified
  by a column in \reftab{countnecklace}.
%

@@

\begin{theorem}
The number $\NB(n)$ of the square-free cyclic polynomials $C_p(\vx)$ ($p \in \Bset$)
  formed by $\vx = \{x_1, \ldots, x_n\}$ is
\begin{equation}
  \NB(n) = N(n) = (1/n) \sum_{d|n} \varphi(n/d) 2^{\,d},
\label{eq:necklace}
\end{equation}
%
which is also equal to $\deg_\lambda A_n\RX$.
\label{thm:necklace}
\end{theorem}
%
$N(n)$ = 2, 3, 4, 6, 8, 14, 20, 36, 60, 108, 188, 352, 632, \ldots,
  starting from $n = 1$.

@@

Since the period $d$ of a length-$n$ necklace always divides $n$,
we have
\begin{equation}
N(n) = \sum_{d|n} L(d).
\label{eq:necklacelyndon}
\end{equation}
%
We can also show this by explicit computation:
\begin{align*}
\sum_{d | n} L(d)
  &= \sum_{d | n} (1/d) \sum_{c |  d} \mu(d/c) 2^{\,c}
   = (1/n) \sum_{c | n} 2^{\,c} \sum_{d'  |  (n/c)}
        \mu\left(\frac{n/c}{d'}\right) d' \\
  &= (1/n) \sum_{c \, | \, n} 2^{\,c} \, \varphi(n/c)
  = N(n),
\end{align*}
where we have used $\varphi(m) = \sum_{d' | m} \mu(\frac{m}{d'}) \, d'$,
which is the inversion of $m = \sum_{d' | m} \varphi(d')$,
and \refeq{necklace}.
%
We will use \refeq{necklacelyndon} in proving the next theorem.

@@

\begin{theorem}
The degrees in $R$ of $A_n\RX$, $P_n\RX$ and $S_n(R)$ are
\begin{subequations}
\begin{align}
\deg_R A_n\RX &= \sum_{d|n} \varphi(n/d) 2^{d-1}, \\
\deg_R P_n\RX &= \sum_{d|n} \mu(n/d) 2^{d-1} \equiv \beta(n), \\
\deg S_n(R) &= \beta(n) - \sum_{d | n, d < n} \beta(d) \, \varphi(n/d).
\end{align}
\label{eq:degR}
\end{subequations}
\label{thm:degR}
\end{theorem}

@@

\begin{figure}[h]
  \begin{minipage}{\linewidth}
    \begin{center}
        \includegraphics[angle=-90, width=\linewidth]{cubbulb.pdf}
    \end{center}
  \end{minipage}%
  \caption{
  \label{fig:cubbulb}
  Stable regions of the $n$-cycles of the cubic map \refeq{cubic}
  with a complex $R$;
  obtained from $P_n(R, e^{i\theta}) = 0$ with $\theta \in [0, 2\pi]$
  cf. \reffig{logbulb}.
}
\end{figure}

@@

%Now, the number of the cube-free cyclic polynomials
%is now equal to that of total necklaces
%\[
%  N_3(n) = (1/n) \sum_{d|n} \varphi(n/d) 3^d.
%\]
%
%The proof is similar to that in \refsec{degprimfac}.
%For even necklaces, we have

\begin{theorem}
The number of the even ternary necklaces
or the cube-free cyclic polynomials for the cubic map of even degrees in $\vx$
is given by
\begin{equation}
  N_e(n) = \frac{1}{n} \sum_{c d = n} \varphi(c)
    \left[
      3^d - \odd(c) \frac{3^d-1}{2}
    \right],
\label{eq:cubnecklace}
\end{equation}
where
$\odd(c) \equiv \big[1 - (-1)^c\big]/2$ is 1 if $c$ is odd or 0 if even.
\label{thm:cubnecklace}
\end{theorem}

@@

\begin{table}[h]\footnotesize
\tbl{
%\caption{
Two ways of counting the $N_e(2) = 4$ ternary necklaces for $n = 2$.
}
%\begin{center}
{
\begin{tabular}{c | c  c c | c  | c | c | c | c }
%\begin{tabularx}{\textwidth}{c | c  c c | c  | c | c | c | c }
\hline
        & \multicolumn{3}{c|}{$m = 1$}
        & $m = 2$
        & $c = n/d$
        & $T(d, c)$ $^\dagger$
        & $\varphi(c)$ & $\varphi(n/d) T(d, c)$\\
\hline
$d = 1$ & \hspace{1mm} $\mathbf{0}_{\times2}$ \hspace{1mm}
        & \hspace{1mm} $\mathbf{1}_{\times2}$ \hspace{1mm}
        & \hspace{1mm} $\mathbf{2}_{\times2}$ \hspace{1mm}
        & & $2$ & $3$ & $1$ & 3 \\
\hline
$d = 2$ & $\mathbf{00}$
        & $\mathbf{11}$
        & $\mathbf{22}$
        & \hspace{1mm} $\mathbf{20} (02)$ $^\ddagger$ \hspace{1mm}
        & $1$ & $3+1\cdot2=5$ & $1$ & 5 \\
\hline
$m \sum_{m|d, d|n} \varphi(\frac{n}{d})$
      & \multicolumn{3}{c|}{$1\cdot(1+1) = 2$}
      & $2\cdot1 = 2$ &    \multicolumn{3}{c|}{} &  $\downarrow$\\
\hline
$N_e(n)\cdot n$
    & \multicolumn{3}{c|}{$3 \cdot 2$} & $1 \cdot 2$ &
  \multicolumn{3}{c|}{$\rightarrow$} & $4\cdot2 = 3 + 5$ \\
\hline
\multicolumn{9}{p{\linewidth}}{
  Bold strings are necklaces; others are their cyclic versions.
}
\\
\multicolumn{9}{p{\linewidth}}{
The subscript of a string means the number of repeats;
e.g.,
$\mathbf{1}_{\times2}$ means $\mathbf{1}$ repeated twice, or $\mathbf{11}$;
}
\\
\multicolumn{9}{p{\linewidth}}{
$^\dagger$ $T(d, c) \equiv 3^d - \odd(c)(3^d-1)/2$,
which is $3^d$ if $c$ is even,
or $(3^d+1)/2$ if $c$ is odd.
}
\\
\multicolumn{9}{p{\linewidth}}{
$^\ddagger$
The odd binary strings $\mathbf{10}$, $01$, $\mathbf{21}$, and $12$
  do not contribute to the sum, and are excluded from $T(d, c)$.
}
\\
\hline
\end{tabular}
%\end{tabularx}
%\end{center}
\label{tab:countcubnecklace}
}
\end{table}

The rest counting process is similar to that in \refsec{degprimfac}:
%
we construct the sum $n N_e(n)$ by two ways, as exemplified in \reftab{countcubnecklace}.
%
In the first way,
  for a fixed $d$ ($d|n$),
  if $c\equiv n/d$ is even,
    we count all ternary strings whose period $m$ divide $d$,
  but if $c$ is odd, we count only ternary strings
    whose first $d$ beads are even
    (because repeating an odd string an odd number times
    does not yield an even string);
  in either case, we multiple the result by $\varphi(c)$.
%
The process for a fixed $d$ corresponds to a row of \reftab{countcubnecklace}.
%
Repeating the process for all $d$ gives
  $\sum_{cd = n} T(d, c) \, \varphi(c)$,
  where $T(d, c)$ is the total number $3^d$ of period-$d$ ternary strings if $c$ is even,
  or the number $(3^d+1)/2$ of even strings if $c$ is odd.

In the second way,
  we look at the contribution from each necklace to the above sum.
An even period-$m$ necklace contributes a total of
  $m \times \sum_{m|d, d|n} \varphi(n/d) = n$
  \big[the multiplier $m$ is for the $m$ cyclic versions, cf. \refeq{mphind}\big],
  while an odd necklace contributes nothing.
Thus, the sum equals $N_e(n)\cdot n$.
%
The process for a fixed necklace corresponds to
  a column of \reftab{countcubnecklace}.
